Consistent Polyhedral Surrogates via Embedding

### 2019-06-20 notes/raf-lovasz.tex
Jessie: Added comments and questions on the Lovasz writeup

A natural approach to deriving convex surrogates in the case of submodular functions $f$ is the Lov\'asz extension, which is known to be convex when $f$ if (and only if) submodular.
Given any set-valued function $f:2^N\to\reals$, its \emph{Lov\'asz extension} $F:\reals^k\to\reals$ is given by
\begin{equation}\label{eq:lovasz-ext}
  F(u) = \E[f(\{i:u_i \geq \Theta\})]~,
% F(u) = \sum_{i=1}^k u_{j_i} (f(\{j_1,\ldots,j_i\}) - f(\{j_1,\ldots,j_{i-1}\})~,
\end{equation}
where $\Theta$ is a uniformly distributed random variable on $[0,1]$.
There are several equivalent formulations for the Lov\'asz extension; see~\citet[Definition 3.1]{bach2013learning}.

@@

  In the other case, $(\chi_S)_iu_i < 0$, we have $w_i = 1+|u_i| > 2$ and $w'_i = 1+|u'_i| = 2$, and again the other entries are identical.
  In particular, $w'_i \leq w_i$.
  Moreover, we claim that there is other value in between, i.e.,\ there is no index $j\neq i$ such that $w'_i < w_j < w_i$.
  This follows from our assumption on $i$: if we had such a $j$, then we must have $2 < 1 - (\chi_S)_ju_j < 1 + |u_i|$, which can only occur when $\sgn(u_j) \neq \chi_S$, and thus $-(\chi_S)_ju_j = |u_j|$; we conclude $1 < |u_j| < |u_i|$, a contradiction.
  Thus, for all $j\neq i$, we have either $w_j \leq w'_i \leq w_i$ or $w'_i \leq w_i \leq w_j$.
  In light of $w'_j = w_j$, this is equivalent to either (a) $w_j \leq w_i$ and $w'_j \leq w'_i$, or (b) $w_i \leq w_j$ and $w'_i \leq w'_j$.
  Thus, there is a permutation $\pi$ which orders the elements of both $w$ and $w'$ simultaneously: for all $j,j'\in [k]$, $j<j'$, we have both $w_{\pi_j} \geq w_{\pi_{j'}}$ and $w'_{\pi_j} \geq w'_{\pi_{j'}}$.
  By another common representation of the Lov\'asz extension~\cite[Equation 3.1]{bach2013learning}, we thus have $F(w_+) - F(w'_+) = ((w_+)_i - (w'_+)_i)(f(T\cup\{i\})-f(T)) = (|u_i|-1)(f(T\cup\{i\})-f(T)) > 0$,
  where $T = \{\pi_1,\ldots,\pi_j\}$ such that $\pi_{j+1} = i$, and we have used the fact that $f$ is increasing and $|u_i|>1$.
\end{proof}

@@

  Condition (i) follows immediately: for any $u\in[-1,1]^k$, let $A = \{i:u_i\geq 0\}$, and $\pi$ be any permutation ordering the elements of $u \odot \chi_A$.
  For condition (ii), note that for any $u\in P^A_\pi$, as $u\odot \chi_A \in P_\pi$ we may write
  \begin{equation}
    \label{eq:udot-conv}
    u \odot \chi_A = \sum_{i=1}^{k-1} \left( u_{\pi_i} (\chi_A)_{\pi_i} - u_{\pi_{i+1}} (\chi_A)_{\pi_{i+1}} \right) \ones_{\{\pi_1,\ldots,\pi_i\}} + u_{\pi_k} (\chi_A)_{\pi_k} \ones
    % + (1-u_{\pi_1}(\chi_A)_{\pi_1} \cdot 0
    ~,
  \end{equation}
  which is a convex combination (again, see \citet[pg.\ 167]{bach2013learning}).
  We simply apply the involution $\odot\chi_A$ again, to obtain
  \begin{equation}
    \label{eq:udot-conv}
    u = \sum_{i=1}^{k-1} \left( u_{\pi_i} (\chi_A)_{\pi_i} - u_{\pi_{i+1}} (\chi_A)_{\pi_{i+1}} \right) \ones_{\{\pi_1,\ldots,\pi_i\}}\odot \chi_A + u_{\pi_k} (\chi_A)_{\pi_k} \chi_A
    % + (1-u_{\pi_1}(\chi_A)_{\pi_1} \cdot 0
    ~,
  \end{equation}
  and condition (ii) follows as $\ones_B \cdot \chi_A \in \U$ for all sets $B\subseteq N$.

@@

  To construct $\pi^S$, we ``shatter'' the permutation $\pi$ into two pieces, depending on whether an index is in $A\triangle S$ or not.
  In particular, note that if $i\in A\triangle S$, then for all $u\in P^A_\pi$ we have $u_i(\chi_A)_i \geq 0$ and $(\chi_A)_i = -(\chi_S)_i$, so $(\ones - u\odot\chi_S)_i = 1 - u_i(\chi_S)_i = 1 + u_i(\chi_A)_i \geq 1$.
  Similarly, when $i\notin A\triangle S$, then $(\chi_A)_i = (\chi_S)_i$, so $(\ones - u\odot\chi_S)_i = 1 - u_i(\chi_S)_i = 1 - u_i(\chi_A)_i \leq 1$.
  As $\pi$ orders the elements $u_i(\chi_S)_i$ in decreasing order, we see that the following permutation $\pi^S$ will order the elements of $\ones - u\odot\chi_S$ in decreasing order: order the elements in $A\triangle S$ according to $\pi$, followed be the remaining elements according to the reverse of $\pi$.
  As the definition of $\pi^S$ is independent of $u$, we see that $u\in P^A_\pi \implies \ones - u\odot\chi_S \in P_{\pi^S}$, as desired.
\end{proof}

@@

\begin{lemma}
  \label{lem:lovasz-u}
  Let $\hat\ell^f:\hat\R\times\Y\to\reals_+$, where $\hat\R = \{(A,B) \in 2^N \times 2^N: A\cap B=\emptyset\}$, be given by
  \begin{equation}
    \label{eq:lovasz-embeds}
    \hat\ell^f((A,B),S) = L^f(\chi_A + \ones_B,S) = f(A\triangle S\setminus B) + f(A\triangle S\cup B)~.
  \end{equation}
  Then the Lov\'asz hinge $L^f$ embeds $\hat\ell^f$.
\end{lemma}
\begin{proof}
  As observed above, the set $\U$ is in bijection with $\hat \R$, using the transformation $u = \chi_A + \ones_B$.
  As also observed above, we may write $u = \chi_A \odot \ones_{N\setminus B}$ as well.
  Combining Lemma~\ref{lem:lovasz-trim} with \raf{RESTRICTION LEMMA}, we see that $L^f$ embeds $L^f|_\U$.
  It thus only remains to verify the form~\eqref{eq:lovasz-embeds}.
  We have $L^f(u,S) = F(x)$ where $x = \ones - u\odot\chi_S = \ones - \ones_{N\setminus B} \odot \chi_A \odot \chi_S = \ones + \ones_{N\setminus B} \odot \chi_{A\triangle S} = \ones_{B} + 2\ones_{A\triangle S\setminus B}$.
  As $(A\triangle S \setminus B) \cup B = A\triangle S \cup B$, the result follows from~\cite[Prop 3.1(h)]{bach2013learning}.
\end{proof}

@@

\begin{lemma}
  \label{lem:2-bar-f}
  For all $(A,B) \in \hat\R$, $\hat\ell^f((A,B);\bar p) \geq f(N)$.
  For all $A\subseteq N$, $\hat\ell^f((A,\emptyset);\bar p) = 2\bar f$.
\end{lemma}
\begin{proof}
  Letting $\overline B := N\setminus B$ for short, we have
  \begin{align*}
    \hat\ell^f((A,B);\bar p)
    &= 2^{-k} \sum_{S\subseteq N} f(A\triangle S\setminus B) + f(A\triangle S\cup B)
    \\
    &= 2^{-|\overline B|} \sum_{T\subseteq \overline B} f(T) + f(T\cup B)
    \\
    &= \frac 1 2 \; 2^{-|\overline B|} \sum_{T\subseteq \overline B} f(T) + f(\overline B\setminus T) + f(T\cup B) + f((\overline B\setminus T)\cup B)
    \\
    &\geq \frac 1 2 \left( f(\overline B) + f(\emptyset) + f(N) + f(B) \right)
    \\
    &\geq \frac 1 2 \left( f(N) + f(N) \right) = f(N)~,
  \end{align*}
  where we use submodularity in both inequalities.
  The second statement follows from the second equality above after setting $B=\emptyset$, as then $\overline B = N$ and thus $T$ ranges over all of $2^N$.
\end{proof}

@@

  Define $\epsilon = \bar f / (2\bar f - f(N))$, which is strictly positive by Lemma~\ref{lem:bar-f}.
  Let $p = (1-\epsilon) \bar p + \epsilon \delta_\emptyset$, where again $\bar p$ is the uniform distribution, and $\delta_\emptyset$ is the point distribution on $\emptyset$.
  From Lemma~\ref{lem:2-bar-f}, for all $A\subseteq N$ we have
  \begin{align*}
    \hat\ell^f((A,\emptyset);p)
    &= (1-\epsilon) 2 \bar f + \epsilon \, \hat\ell^f((A,\emptyset),\emptyset)\\
    &= (1-\epsilon) 2 \bar f + \epsilon \, 2f(A)\\
    &\geq (1-\epsilon)2 \bar f > f(N) = \hat\ell^f((\emptyset,N);p)~.
  \end{align*}
  As we have some report with strictly lower loss than all reports of the form $(A,\emptyset)$, we conclude that we must have some $(A,B) \in \prop{\hat\ell^f}(p)$ with $B\neq\emptyset$.
  We can also see that $\prop{\ell^f}(p) = \{\emptyset\}$ by the second equality and the fact that $\{\emptyset\} = \argmin_{S\subseteq N} f(S)$.

>>>>

A natural approach to deriving convex surrogates in the case of submodular functions $f$ is the Lov\'asz extension, which is known to be convex when $f$ if (and only if) submodular. \jessiet{Doesn't make grammatical sense to me.}
Given any set-valued function $f:2^N\to\reals$, its \emph{Lov\'asz extension} $F:\reals^k\to\reals$ is given by
\begin{equation}\label{eq:lovasz-ext}
  F(u) = \E[f(\{i:u_i \geq \Theta\})]~,
% F(u) = \sum_{i=1}^k u_{j_i} (f(\{j_1,\ldots,j_i\}) - f(\{j_1,\ldots,j_{i-1}\})~,
\end{equation}
where $\Theta$ is a uniformly distributed random variable on $[0,1]$.
There are several equivalent formulations for the Lov\'asz extension; see~\citet[Definition 3.1]{bach2013learning}.

@@

  In the other case, $(\chi_S)_iu_i < 0$, we have $w_i = 1+|u_i| > 2$ and $w'_i = 1+|u'_i| = 2$, and again the other entries are identical.
  In particular, $w'_i \leq w_i$.
  Moreover, we claim that there is \jessie{ADD: no} other value in between, i.e.,\ there is no index $j\neq i$ such that $w'_i < w_j < w_i$.
  This follows from our assumption on $i$: if we had such a $j$, then we must have $2 < 1 - (\chi_S)_ju_j < 1 + |u_i|$, which can only occur when $\sgn(u_j) \neq \chi_S$, and thus $-(\chi_S)_ju_j = |u_j|$; we conclude $1 < |u_j| < |u_i|$, a contradiction.
  Thus, for all $j\neq i$, we have either $w_j \leq w'_i \leq w_i$ or $w'_i \leq w_i \leq w_j$.
  In light of $w'_j = w_j$, this is equivalent to either (a) $w_j \leq w_i$ and $w'_j \leq w'_i$, or (b) $w_i \leq w_j$ and $w'_i \leq w'_j$.
  Thus, there is a permutation $\pi$ which orders the elements of both $w$ and $w'$ simultaneously: for all $j,j'\in [k]$, $j<j'$, we have both $w_{\pi_j} \geq w_{\pi_{j'}}$ and $w'_{\pi_j} \geq w'_{\pi_{j'}}$.
  By another common representation of the Lov\'asz extension~\cite[Equation 3.1]{bach2013learning}, we thus have $F(w_+) - F(w'_+) = ((w_+)_i - (w'_+)_i)(f(T\cup\{i\})-f(T)) = (|u_i|-1)(f(T\cup\{i\})-f(T)) > 0$,
  where $T = \{\pi_1,\ldots,\pi_j\}$ such that $\pi_{j+1} = i$, and we have used the fact that $f$ is increasing and $|u_i|>1$.
  \jessiet{I think this should technically be a $\geq 0$, since we have no notion of ``strictly increasing'' so even if $f$ is increasing, we could still have $f(T\cup\{i\})-f(T) = 0$.  Doesn't affect the result though.}
\end{proof}

@@

  Condition (i) follows immediately: for any $u\in[-1,1]^k$, let $A = \{i:u_i\geq 0\}$, and $\pi$ be any permutation ordering the elements of $u \odot \chi_A$.
  For condition (ii), note that for any $u\in P^A_\pi$, as $u\odot \chi_A \in P_\pi$ we may write
  \begin{equation}
    \label{eq:udot-conv}
    u \odot \chi_A = \sum_{i=1}^{k-1} \left[ \left( u_{\pi_i} (\chi_A)_{\pi_i} - u_{\pi_{i+1}} (\chi_A)_{\pi_{i+1}} \right) \ones_{\{\pi_1,\ldots,\pi_i\}} \right] + u_{\pi_k} (\chi_A)_{\pi_k} \ones
    % + (1-u_{\pi_1}(\chi_A)_{\pi_1} \cdot 0
    ~,
  \end{equation}
  \jessiet{The indicator notation is confusing to me with the ones vector also in the same context.  That's just a personal opinion though.}
  \jessiet{Unclear to me at first what the summation is over.  Added brackets to clarify.}
  which is a convex combination (again, see \citet[pg.\ 167]{bach2013learning}).
  We simply apply the involution $\odot\chi_A$ again, to obtain
  \begin{equation}
    \label{eq:udot-conv}
    u = \sum_{i=1}^{k-1} \left( u_{\pi_i} (\chi_A)_{\pi_i} - u_{\pi_{i+1}} (\chi_A)_{\pi_{i+1}} \right) \ones_{\{\pi_1,\ldots,\pi_i\}}\odot \chi_A + u_{\pi_k} (\chi_A)_{\pi_k} \chi_A
    % + (1-u_{\pi_1}(\chi_A)_{\pi_1} \cdot 0
    ~,
  \end{equation}
  and condition (ii) follows as $\ones_B \cdot \chi_A \in \U$ for all sets $B\subseteq N$.
  \jessiet{What is $B$ here? I might just be slow, but I missed the step where this leads to Condition (ii)}

@@

  To construct $\pi^S$, we ``shatter'' the permutation $\pi$ into two pieces, depending on whether an index is in $A\triangle S$ or not.
  In particular, note that if $i\in A\triangle S$, then for all $u\in P^A_\pi$ we have $u_i(\chi_A)_i \geq 0$ and $(\chi_A)_i = -(\chi_S)_i$, so $(\ones - u\odot\chi_S)_i = 1 - u_i(\chi_S)_i = 1 + u_i(\chi_A)_i \geq 1$.
  Similarly, when $i\notin A\triangle S$, then $(\chi_A)_i = (\chi_S)_i$, so $(\ones - u\odot\chi_S)_i = 1 - u_i(\chi_S)_i = 1 - u_i(\chi_A)_i \leq 1$.
  As $\pi$ orders the elements $u_i(\chi_S)_i$ in decreasing order, we see that the following permutation $\pi^S$ will order the elements of $\ones - u\odot\chi_S$ in decreasing order: sort the elements in $A\triangle S$ according to $\pi$, followed by the remaining elements according to the reverse of $\pi$.
  As the definition of $\pi^S$ is independent of $u$, we see that $u\in P^A_\pi \implies \ones - u\odot\chi_S \in P_{\pi^S}$, as desired.
\end{proof}

@@

\begin{lemma}
  \label{lem:lovasz-u}
  Let $\hat\ell^f:\hat\R\times\Y\to\reals_+$, where $\hat\R = \{(A,B) \in 2^N \times 2^N: A\cap B=\emptyset\}$, be given by
  \begin{equation}
    \label{eq:lovasz-embeds}
    \hat\ell^f((A,B),S) = L^f(\chi_A + \ones_B,S) = f(A\triangle S\setminus B) + f(A\triangle S\cup B)~.
  \end{equation}
  Then the Lov\'asz hinge $L^f$ embeds $\hat\ell^f$.
\end{lemma}
\begin{proof}
  As observed above, the set $\U$ is in bijection with $\hat \R$, using the transformation $u = \chi_A + \ones_B$.
  As also observed above, we may write $u = \chi_A \odot \ones_{N\setminus B}$ as well.
  Combining Lemma~\ref{lem:lovasz-trim} with \raf{RESTRICTION LEMMA}, we see that $L^f$ embeds $L^f|_\U$.
  It thus only remains to verify the form~\eqref{eq:lovasz-embeds}.
  We have $L^f(u,S) = F(x)$ where $x = \ones - u\odot\chi_S = \ones - \ones_{N\setminus B} \odot \chi_A \odot \chi_S = \ones + \ones_{N\setminus B} \odot \chi_{A\triangle S} = \ones_{B} + 2\ones_{A\triangle S\setminus B}$.
  \jessiet{This next line lost me.}
  As $(A\triangle S \setminus B) \cup B = A\triangle S \cup B$, the result follows from~\cite[Prop 3.1(h)]{bach2013learning}.
\end{proof}

@@

\begin{lemma}
  \label{lem:2-bar-f}
  For all $(A,B) \in \hat\R$, $\hat\ell^f((A,B);\bar p) \geq f(N)$. \jessiet{Intuitively, this seems backwards?  Since $f$ is increasing.  But that might be the point.}
  For all $A\subseteq N$, $\hat\ell^f((A,\emptyset);\bar p) = 2\bar f$.
\end{lemma}
\begin{proof}
  Letting $\overline B := N\setminus B$ for short \jessiet{To make notation more consistent, maybe change $\overline B$ to $B^c$ so something else to deviate from the average notation.}, we have
  \begin{align*}
    \hat\ell^f((A,B);\bar p)
    &= 2^{-k} \sum_{S\subseteq N} f(A\triangle S\setminus B) + f(A\triangle S\cup B)
    \\
    &= 2^{-|\overline B|} \sum_{T\subseteq \overline B} f(T) + f(T\cup B)
    \\
    &= \frac 1 2 \; 2^{-|\overline B|} \sum_{T\subseteq \overline B} f(T) + f(\overline B\setminus T) + f(T\cup B) + f((\overline B\setminus T)\cup B)
    \\
    &\geq \frac 1 2 \left( f(\overline B) + f(\emptyset) + f(N) + f(B) \right)
    \\
    &\geq \frac 1 2 \left( f(N) + f(N) \right) = f(N)~,
  \end{align*}
  where we use submodularity in both inequalities.
  \jessiet{I didn't see how we got from first line to second until this little reminder that $B = \emptyset$ below-- any chance we can move that above the arithmetic?}
  The second statement follows from the second equality above after setting $B=\emptyset$, as then $\overline B = N$ and thus $T$ ranges over all of $2^N$.
\end{proof}

@@

  Define $\epsilon = \bar f / (2\bar f - f(N))$, which is strictly positive by Lemma~\ref{lem:bar-f} \jessie{and submodularity of $f$}.
  Let $p = (1-\epsilon) \bar p + \epsilon \delta_\emptyset$, where again $\bar p$ is the uniform distribution, and $\delta_\emptyset$ is the point distribution on $\emptyset$.
  From Lemma~\ref{lem:2-bar-f}, for all $A\subseteq N$ we have
  \begin{align*}
    \hat\ell^f((A,\emptyset);p)
    &= (1-\epsilon) 2 \bar f + \epsilon \, \hat\ell^f((A,\emptyset),\emptyset)\\
    &= (1-\epsilon) 2 \bar f + \epsilon \, 2f(A)\\
    &\geq (1-\epsilon)2 \bar f > f(N) = \hat\ell^f((\emptyset,N);p)~.
  \end{align*}
  As we have some report with strictly lower loss than all reports of the form $(A,\emptyset)$, we conclude that we must have some $(A,B) \in \prop{\hat\ell^f}(p)$ with $B\neq\emptyset$.
  We can also see that $\prop{\ell^f}(p) = \{\emptyset\}$ by the second equality and the fact that $\{\emptyset\} = \argmin_{S\subseteq N} f(S)$.
